## A First Example
```{r echo = FALSE, message = FALSE, warning = FALSE}
library(Rcpp)
library(inline)
setwd("~/Documents/GitHub/Seamless_R_and_Cpp/")
```

The Fibonacci sequence $F_n$ is defined by $F_n = F_{n-1} + F_{n-2}$ with 2 initial conditions $F_0 = 0$ and $F_1 = 1$. 

### R solution
```{r}
fibR <- function(n) {
  if (n == 0) return(0)
  if (n == 1) return(1)
  return (fibR(n - 1) + fibR(n - 2))
}

```

### C++ solution
```{r}
# require Rcpp, inline
incltxt <- '
int fibonacci(const int x) {
  if (x == 0) return(0);
  if (x == 1) return(1);
  return fibonacci(x - 1) + fibonacci(x - 2);
}
'

fibRcpp <- cxxfunction(signature(xs="int"),
                       plugin="Rcpp",
                       incl=incltxt,
                       body='
         int x = Rcpp::as<int>(xs) ;
         return Rcpp::wrap( fibonacci(x) ) ;
         ')
```

> Note: This uses two key Rcpp tools, the converter functions `as` and `wrap`. 
> 
> 1. `as` is used to convert the incoming argument `xs` from `SEXP` to integer. 
> 2. `wrap` converts the integer result in the integer variable `x` to the `SEXP` type. 
> 3. In `incltxt`, we define a C++ function `fibonacci`. 
> 4. `cxxfunction`: is used to turn code supplied as a text variable into an executable function.

```{r}
rbenchmark::benchmark(fibR(20), fibRcpp(20))
```

### Using Rcpp Attributes
One can simply declare certain “attributes,” notably whether a function is to be exported for use from R or from another C++ function (or both). One can declare dependencies whose resolution still relies on the plugin framework provided by inline.
```{r}
# in c1.2.cpp: 
# #include <Rcpp.h>
# using namespace Rcpp;
# 
# // [[Rcpp::export]]
# int fibonacci(const int x) {
#   if (x< 2)
#     return x;
#   else
#     return (fibonacci(x - 1)) + fibonacci(x - 2);
# }

sourceCpp("c++/c1.2.cpp")
fibonacci(20)
```

> Note: 
> 
> 1. The `[[Rcpp::export]]` attribute preceds the function definition.
> 2. The `sourceCpp()` reads the code from the given source file, parses it for the relevant attributes, and creates the required wrappers before calling R to compile and link just like `inline` does.



