## A First Example
```{r echo = FALSE, message = FALSE, warning = FALSE}
library(Rcpp)
library(inline)
setwd("~/Documents/GitHub/Seamless_R_and_Cpp/")
```

The Fibonacci sequence $F_n$ is defined by $F_n = F_{n-1} + F_{n-2}$ with 2 initial conditions $F_0 = 0$ and $F_1 = 1$. 

### R solution {-}
```{r}
fibR <- function(n) {
  if (n == 0) return(0)
  if (n == 1) return(1)
  return (fibR(n - 1) + fibR(n - 2))
}

```

### C++ solution {-}
```{r}
# require Rcpp, inline
incltxt <- '
int fibonacci(const int x) {
  if (x == 0) return(0);
  if (x == 1) return(1);
  return fibonacci(x - 1) + fibonacci(x - 2);
}
'

fibRcpp <- cxxfunction(signature(xs="int"),
                       plugin="Rcpp",
                       incl=incltxt,
                       body='
         int x = Rcpp::as<int>(xs) ;
         return Rcpp::wrap( fibonacci(x) ) ;
         ')
```

> Note: This uses two key Rcpp tools, the converter functions `as` and `wrap`. 
> 
> 1. `as` is used to convert the incoming argument `xs` from `SEXP` to integer. 
> 2. `wrap` converts the integer result in the integer variable `x` to the `SEXP` type. 
> 3. In `incltxt`, we define a C++ function `fibonacci`. 
> 4. `cxxfunction`: is used to turn code supplied as a text variable into an executable function.

```{r}
rbenchmark::benchmark(fibR(20), fibRcpp(20))
```

### Using Rcpp Attributes {-}
One can simply declare certain “attributes,” notably whether a function is to be exported for use from R or from another C++ function (or both). One can declare dependencies whose resolution still relies on the plugin framework provided by inline.
```{r}
# in c1.2.cpp: 
# #include <Rcpp.h>
# using namespace Rcpp;
# 
# // [[Rcpp::export]]
# int fibonacci(const int x) {
#   if (x< 2)
#     return x;
#   else
#     return (fibonacci(x - 1)) + fibonacci(x - 2);
# }

sourceCpp("c++/c1.2.cpp")
fibonacci(20)
```

> Note: 
> 
> 1. The `[[Rcpp::export]]` attribute preceds the function definition.
> 2. The `sourceCpp()` reads the code from the given source file, parses it for the relevant attributes, and creates the required wrappers before calling R to compile and link just like `inline` does.

### A second R solution {-}
One elegant solution to retain the basic recursive structure of the algorithm with- out incurring the cost of repeated computation of the same value is provided by a method called **memoization**. 
```{r}
mfibR <- local({
  memo <- c(1, 1, rep(NA, 1000)) 
  f <- function(x) {
    if (x == 0) return(0) 
    if (x < 0) return(NA)
    if (x > length(memo))
      stop("x too big for implementation") 
    if (!is.na(memo[x])) return(memo[x])
    ans <- f(x-2) + f(x-1)
    memo[x] <<- ans
    ans
    } 
  })
```

If a value for argument n has already been encountered, it is used. Otherwise, it is computed and stored in vector memo. This ensures that the recursive function is called exactly once for each possible value of n. 

> Note: 
> 
> 1. `local` function and `<<-` 
> 2. run `mfibR(x)`

### A second C++ solution {-}
```{r}
mincltxt<- '
#include <algorithm>
#include <vector>
#include <stdexcept>
#include <cmath>
#include <iostream>

class Fib {
public:
  Fib(unsigned int n = 1000) {
    memo.resize(n);     // reserve n elements
    std::fill( memo.begin(), memo.end(), NAN );     // set to NaN
    memo[0] = 0.0;      // initialize for
    memo[1] = 1.0;      // n=0 and n=1
  }
  double fibonacci(int x)  {
    if (x < 0)   // guard against bad input
      return ( (double) NAN );
    if (x >= (int) memo.size() )
      throw std::range_error(\"x too large for implementation\");
    if ( ! ::isnan(memo[x]))
      return (memo[x]) ;   // if exist, reuse values
    // build precomputed value via recursion
    memo[x] = fibonacci(x-2) + fibonacci(x-1) ;
    return ( memo[x] ) ;    // and return
  }
private:
  std::vector< double > memo ;  // internal memory for precomp.
};
'
mfibRcpp <- cxxfunction(signature(xs="int"),
                        plugin="Rcpp", includes=mincltxt,
                        body='
              int x = Rcpp::as<int>(xs);
              Fib f;
              return Rcpp::wrap( f.fibonacci(x-1) );
')
```

We define a very simple C++ class Fib with three elements:
+ A constructor which is called once upon initialization.
+ A single public member function which computes Fn.
+ A private data vector holding the **memoization** values.


```{r}
rbenchmark::benchmark(fibRcpp(30), mfibR(30), mfibRcpp(30))
```

### A third R solution {-}
We can also compute $F_n$ using iterative approach. The iterative solution improves further on the approach using memoization as it requires neither stateful memory nor recursion. 
```{r}
fibRiter<-function(n){
  first <- 0
  second <- 1
  third <- 0
  for (i in seq_len(n)) {
    third <- first + second 
    first <- second
    second <- third
  }
  return(first)
}
```


### A third C++ solution {-}
```{r}
fibRcppIter<-cxxfunction(signature(xs="int"), plugin="Rcpp",
                         body='
   int n = Rcpp::as<int>(xs);
   double first = 0; 
   double second = 1; 
   double third = 0; 
   for (int i=0; i<n; i++) {
    third = first + second; 
    first = second;
    second = third;
   }
  return Rcpp::wrap(first);
')
```

> Note: 
> 
> 1. From this example, we do not need to write a Cpp function separately. We can put the Cpp function in the `body` argument of `cxxfunction`. 
```{r}
rbenchmark::benchmark(mfibR(100), mfibRcpp(100), fibRiter(100), fibRcppIter(100))
```

